Parsing Postfix log files

### 2007-06-22 doc/logparser.tex
Wrote overlapping rules section.

>>>>

\subsection{Overlapping rules}

The parser does not try to detect overlapping rules;\footnote{It may be
possible to parse each rule's regex, generate an NFA, and compare all NFAs
to determine if any overlap.  The author has not attempted to do this: such
a project would probably qualify for a PhD, and may involve solving the
Halting Problem and circumventing the Church-Turing Thesis.} that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the order in which rules are tried against each line determines which rule
matches.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some lines and a more general rule to match the
remainder, e.g. separating smtp delivery to a filtering proxy from smtp
delivery to the rest of the world.  The algorithm provides a facility for
handling overlapping rules: the priority field in each rule (defaults to
zero).  Rules are tried in order of priority, highest first, overriding the
rule\_order field maintained by the parser.  Negative priorities may also
be useful for catchall rules.

Detecting overlapping rules is difficult, but the following may be helpful:

    \item Sort the regexs and visually inspect the list, e.g. with SQL
        similar to: \textbf{select regex from rules order by regex;}

    \item Compare the results of parsing using sorted, shuffled and
        reversed rules.\footnote{See section~\ref{rule efficiency} for more
        details of sorting the rules.}  Parse a number of log files using
        normal sorting, dump a textual representation of the connections
        and results tables, and delete everything from those tables.
        Repeat with shuffled and reversed sorting.  If there are no
        overlapping rules the tables from each run will be identical;
        differences indicate overlapping rules, and the rule\_ids which
        differ determine which rules overlap.  Unfortunately this method
        cannot prove the absence of overlapping rules; it can detect
        overlapping rules --- but only if there are log lines which match
        more than one rule.
